
```{r setup, include=FALSE}
# title: "hexagon crime incident counts by time period"
# author: "Matthew Ohara"
# date: "2025-04-29"
# revised 5-4-2025 by MK
knitr::opts_chunk$set(echo = TRUE)
```

This function finds the crime count per week per hexagon, then creates the corresponding panel df. In order to ensure that the first and last weeks account for crimes in full weeks (no partial weeks), the function shaves a few days off the beginning and end of the crime data when the crime data does not start with Sunday or end with Saturday. A similar method is used for the create_month_panel function.

```{r}
# first_sunday is the first sunday in the crime data, which represents start of the week
# last_saturday is the last saturday in the crime data, which represents end of the month
# for wday(), the indexes are: sun = 1, mon = 2, .... , sat = 7

create_week_panel <- function(crime, hex) {
  
  # finding first sunday and last saturday of the whole crime dataset
  first_sunday <- as.Date(min(crime$date)) + (7 - wday(as.Date(min(crime$date))) + 1) %% 7
  last_saturday <- max(crime$date) - (wday(max(crime$date)) %% 7)
  
  crime_chopped <- crime %>%
    filter(date >= first_sunday, date <= last_saturday) %>%
    mutate(week = floor_date(date, unit = "week", week_start = 7))
  
  date_range = range(crime_chopped$week)
  
  joined <- st_join(crime_chopped, hex)
  
  # Count of crimes per grid_id per week
  crime_counts_wk <- joined %>%
    st_drop_geometry() %>%
    count(grid_id, week, name = "n_incidents_wk")
  
  all_combos <- expand.grid(
    grid_id = unique(hex$grid_id),
    week = seq(date_range[1], date_range[2], by = "week"),
  stringsAsFactors = FALSE)
  
  panel <- all_combos %>%
    left_join(crime_counts_wk, by = c("grid_id", "week")) %>%
    mutate(n_incidents_wk = replace_na(n_incidents_wk, 0)) %>%
   left_join(hex, by = "grid_id") %>%
    st_as_sf()
  
  return(panel)
}

panel_sc_wk = create_week_panel(df_sc_sm, sc_hex)
panel_sj_wk = create_week_panel(df_sj_sm, sj_hex)
panel_sf_wk = create_week_panel(df_sf_sm, sf_hex)
panel_oak_wk = create_week_panel(df_oak_sm, oak_hex)
panel_la_wk = create_week_panel(df_la_sm, la_hex)

# saveRDS(panel_sc_wk, file=paste0(scrap, "/panel_sc_wk.Rds"))
# saveRDS(panel_sj_wk, file=paste0(scrap, "/panel_sj_wk.Rds"))
# saveRDS(panel_sf_wk, file=paste0(scrap, "/panel_sf_wk.Rds"))
# saveRDS(panel_oak_wk, file=paste0(scrap, "/panel_oak_wk.Rds"))
# saveRDS(panel_la_wk, file=paste0(scrap, "/panel_la_wk.Rds"))
# rm(panel_sc_wk,panel_sj_wk)
# rm(panel_sf_wk,panel_oak_wk, panel_la_wk)
#rm(df_la, df_sc, df_sj, df_sf, df_oak)
```


```{r}

create_day_panel <- function(crime, hex) {
  date_range = range(crime$date, na.rm = TRUE)
  joined <- st_join(crime, hex)
  
  # Count of crimes per grid_id per week
  crime_counts_day <- joined %>%
    st_drop_geometry() %>%
    count(grid_id, date, name = "n_incidents_date")
  
  all_combos <- expand.grid(
    grid_id = unique(hex$grid_id),
    date = seq(date_range[1], date_range[2], by = "day"),
  stringsAsFactors = FALSE)
  
  panel <- all_combos %>%
    left_join(crime_counts_day, by = c("grid_id", "date")) %>%
    mutate(n_incidents_date = replace_na(n_incidents_date, 0)) %>%
   left_join(hex, by = "grid_id") %>%
    st_as_sf()
  
  return(panel)
}
  
panel_sc_day = create_day_panel(df_sc_sm, sc_hex)
panel_sf_day = create_day_panel(df_sf_sm, sf_hex)
panel_oak_day = create_day_panel(df_oak_sm, oak_hex)
panel_la_day = create_day_panel(df_la_sm, la_hex)
#panel_sj_day = create_day_panel(df_sj_sm, sj_hex)

saveRDS(panel_sc_day, file=paste0(scrap, "/panel_sc_day.Rds"))
# saveRDS(panel_sj_day, file=paste0(scrap, "/panel_sj_day.Rds"))
saveRDS(panel_sf_day, file=paste0(scrap, "/panel_sf_day.Rds"))
saveRDS(panel_oak_day, file=paste0(scrap, "/panel_oak_day.Rds"))
saveRDS(panel_la_day, file=paste0(scrap, "/panel_la_day.Rds"))

rm(panel_sc_day,panel_sf_day,panel_oak_day,panel_la_day)
```


This function finds the crime count per month per hexagon, then creates the corresponding panel df.
```{r}
create_month_panel <- function(crime, hex) {
  first_month_start <- floor_date(min(crime$date), unit = "month")
  last_month_end <- ceiling_date(max(crime$date), unit = "month") - days(1)

  crime_chopped = crime %>%
    filter(date >= first_month_start, date <= last_month_end) %>%
    mutate(month_st = floor_date(date, unit = "month"))
  
  date_range <- range(crime_chopped$month_st)
  
  joined <- st_join(crime_chopped, hex)
  
  crime_counts_month <- joined %>%
    st_drop_geometry() %>%
    count(grid_id, month_st, name = "n_incidents_month")
  
  all_combos <- expand.grid(
    grid_id = unique(hex$grid_id),
    month_st = seq(date_range[1], date_range[2], by = "month"),
  stringsAsFactors = FALSE)
  
  panel <- all_combos %>%
    left_join(crime_counts_month, by = c("grid_id", "month_st")) %>%
    mutate(n_incidents_month = replace_na(n_incidents_month, 0)) %>%
    left_join(hex, by = "grid_id") %>%
    st_as_sf()
  
  return(panel)
}

panel_sc_month = create_month_panel(df_sc_sm, sc_hex)
panel_sj_month = create_month_panel(df_sj_sm, sj_hex)
panel_oak_month = create_month_panel(df_oak_sm, oak_hex)
panel_sf_month = create_month_panel(df_sf_sm, sf_hex)
panel_la_month = create_month_panel(df_la_sm, la_hex)

# saveRDS(panel_sc_month, file=paste0(scrap, "/panel_sc_month.Rds"))
# saveRDS(panel_sj_month, file=paste0(scrap, "/panel_sj_month.Rds"))
# saveRDS(panel_oak_month, file=paste0(scrap, "/panel_oak_month.Rds"))
# saveRDS(panel_sf_month, file=paste0(scrap, "/panel_sf_month.Rds"))
# saveRDS(panel_la_month, file=paste0(scrap, "/panel_la_month.Rds"))

rm(panel_sc_month,panel_sj_month,panel_oak_month,panel_sf_month,panel_la_month)

```


```{r}
create_year_panel <- function(crime, hex) {
  first_year_start <- floor_date(min(crime$date), unit = "year")
  last_year_end <- ceiling_date(max(crime$date), unit = "year") - days(1)
  
  crime_chopped <- crime %>%
    filter(date >= first_year_start, date <= last_year_end) %>%
    mutate(year_st = floor_date(date, unit = "year"))
  
  date_range <- range(crime_chopped$year_st)
  
  joined <- st_join(crime_chopped, hex)
  
  crime_counts_year <- joined %>%
    st_drop_geometry() %>%
    count(grid_id, year_st, name = "n_incidents_year")
  
  all_combos <- expand.grid(
    grid_id = unique(hex$grid_id),
    year_st = seq(date_range[1], date_range[2], by = "year"),
    stringsAsFactors = FALSE
  )
  
  panel <- all_combos %>%
    left_join(crime_counts_year, by = c("grid_id", "year_st")) %>%
    mutate(n_incidents_year = replace_na(n_incidents_year, 0)) %>%
    left_join(hex, by = "grid_id") %>%
    st_as_sf()
  
  return(panel)
}

panel_sc_year = create_year_panel(df_sc_sm, sc_hex)
panel_sj_year = create_year_panel(df_sj_sm, sj_hex)
panel_oak_year = create_year_panel(df_oak_sm, oak_hex)
panel_sf_year = create_year_panel(df_sf_sm, sf_hex)
panel_la_year = create_year_panel(df_la_sm, la_hex)

# saveRDS(panel_sc_year, file=paste0(scrap, "/panel_sc_year.Rds"))
# saveRDS(panel_sj_year, file=paste0(scrap, "/panel_sj_year.Rds"))
# saveRDS(panel_oak_year, file=paste0(scrap, "/panel_oak_year.Rds"))
# saveRDS(panel_sf_year, file=paste0(scrap, "/panel_sf_year.Rds"))
# saveRDS(panel_la_year, file=paste0(scrap, "/panel_la_year.Rds"))
# 
# rm(panel_sc_year,panel_sj_year,panel_oak_year,panel_sf_year,panel_la_year)
```

