---
title: "Distance Weighted Shelters"
author: "Jeevan Brar"
date: "`r Sys.Date()`"
output: html_document
---

```{r, message=F, warning=F, comment=NA, echo=T }
# Clear
  rm(list = ls())

# Set working directory
    workdir = getwd()
    setwd(workdir)
library(tidycensus)
library(tidyverse)
library(tigris)
library(sf)
library(tmap)
library(dplyr)
```

## Load Census Tract Shapes and Shelter Locations
We load the shelter data and the ACS tract shapefiles for Alameda, Santa Clara, and San Francisco counties and filter the tracts that fall within the boundaries of Oakland, San Jose, and San Francisco. (note: this code snippet was provided by Jose)

```{r pressure, echo=FALSE}

 shelters_sf <- readRDS(file="/Users/jeevanbrar/Desktop/Econ192/WinterQ/geocoded_shelters_sf.Rdata")
 shelters_scc <- readRDS(file="/Users/jeevanbrar/Desktop/Econ192/WinterQ/geocoded_shelters_scc.Rdata")
  shelters_oak <- readRDS(file="/Users/jeevanbrar/Desktop/Econ192/WinterQ/geocoded_shelters_oak.Rdata")
 # Load ACS variables for 2022 5-year estimates
    v22 <- load_variables(2022, "acs5", cache = TRUE)
# Load ACS tract shapefiles for Alameda, Santa Clara, and San Francisco counties
tract_shapes <- get_acs(
  geography = "tract",
  variables = "B01001_001",
  geometry = TRUE,   
  year = 2022,
  state = "CA",
  county = c("Alameda", "Santa Clara", "San Francisco"),
  output = "wide",   
  survey = "acs5"
) %>% 
  filter(
    !st_is_empty(geometry),
    !GEOID %in% c("06075017903","06075980401")) %>%
  mutate(
    county_fips = substr(GEOID, 1, 5),
    county = recode(county_fips, 
      "06001" = "Alameda",
      "06085" = "Santa Clara",
      "06075" = "San Francisco"
    )
  ) |>
  select(GEOID, county, geometry)

# Get city boundaries for Oakland, San Jose, and San Francisco
city_boundaries <- places(state = "CA", year = 2022, class = "sf") %>%
  filter(NAME %in% c("Oakland", "San Jose", "San Francisco"))

# Filter tracts within the city boundaries
city_shapes <- tract_shapes %>%
  st_filter(city_boundaries)
```


## Convert CRS to be consistent
```{r, message=F, warning=F, comment=NA, echo=T}
shelters <- bind_rows(shelters_sf, shelters_oak, shelters_scc)
# Transform shelters to the same CRS as city_shapes
shelters <- st_transform(shelters, st_crs(city_shapes))

city_shapes <- city_shapes %>% 
  mutate(centroid = st_centroid(geometry))

# Compute distance matrix (shelters × centroids) in meters
distance_matrix <- st_distance(shelters, city_shapes$centroid)

# Find the index of the closest centroid for each shelter
closest_centroid_index <- apply(distance_matrix, 1, which.min)

# Extract the corresponding minimum distance (in meters)
closest_distance <- apply(distance_matrix, 1, min)

# Add the closest centroid and distance as new columns in the shelters dataset
shelters <- shelters %>%
  mutate(
    closest_centroid_id = city_shapes$GEOID[closest_centroid_index],  # Assign the GEOID of the closest tract
    min_distance_meters = closest_distance  # Store the minimum distance
  )


```


## View Census Tracts on Map
```{r, message=F, warning=F, comment=NA, echo=T}


# Create a new sf object with line geometries connecting shelters to closest centroids
shelter_connections <- st_sfc(
  mapply(function(shelter_point, centroid_point) {
    st_linestring(rbind(st_coordinates(shelter_point), st_coordinates(centroid_point)))
  }, st_geometry(shelters), st_geometry(city_shapes$centroid[closest_centroid_index]), SIMPLIFY = FALSE),
  crs = st_crs(city_shapes)  # Ensure same CRS
) %>%
  st_sf(geometry = ., data = shelters %>% select(closest_centroid_id, min_distance_meters))

# Switch to interactive mode
tmap_mode("view")

# Plot census tract boundaries, centroids, shelters, and connections
tm_shape(city_shapes) +  
  tm_borders(lwd = 1, col = "black") +  # Tract boundaries
  
  # Plot centroids
  tm_shape(city_shapes %>% select(centroid)) +  
  tm_dots(col = "red", size = 0.1, alpha = 0.8, title = "Centroids") +  

  # Plot shelters
  tm_shape(shelters) +  
  tm_dots(col = "blue", size = 0.2, alpha = 0.8, title = "Shelters") +  

  # Plot connection lines
  tm_shape(shelter_connections) +  
  tm_lines(lwd = 1, col = "gray", alpha = 0.7) +  

  # Map layout
  tm_layout(title = "Shelter Connections to Closest Census Tract Centroid",
            legend.outside = TRUE)

```

## Define Distance-Weighted Calculation Function and Implement the function
```{r, message=F, warning=F, comment=NA, echo=T}
library(dplyr)

####################################
# 1) Define the new distance-weight
####################################
compute_weight_new <- function(distance_m) {
  # For 1 mile = ~1609 m => discount = 1/(1 + exp(1))
  scale_factor <- 1609
  
  # Single equation for ALL distances (including inside = 0)
  # weight = 1/(1 + exp(distance_m / scale_factor))
  w <- 1 / (1 + exp(distance_m / scale_factor))
  return(w)
}

###########################################
# 2) Apply the new weighting to each shelter
###########################################
shelters <- shelters %>%
  mutate(
    # Compute the new distance-based weight for each shelter
    weight = sapply(min_distance_meters, compute_weight_new),
    
    # Weighted beds is now simply the bed count times the discount factor
    weighted_beds = Total.Beds * weight
  )

#######################################
# 3) Summarize by "closest_centroid_id"
#######################################
exposure_by_tract <- shelters %>%
  group_by(closest_centroid_id) %>%
  summarize(
    exposure = sum(weighted_beds, na.rm = TRUE),
    number_of_shelters = n()
  ) %>%
  ungroup()

# View the results
print(exposure_by_tract)

```


